---
title: "Final_Project_ADA"
author: "Giang Pham"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Loading Packages}
##open libraries
library(haven) #for read sav
library(magrittr) #for piping
library(nnet) #for multinomial logistic regression
library(tidyverse) #for various packages
library(knitr) #for making tables
library(sjlabelled) #show labels on table
library(labelled) #assign names for values
library(ggplot2) #for making plots
library(lmtest) #for lrtest
library(lme4) #comparing coefs
library(gmodels) #for crosstab tables
library(car) #to get vif
library(mice) #check missing values

```

##Setting up the data
```{r Setting Data}
AddHeealth <- read_sav("21600-0022-Data.sav")

Final_Project <- AddHeealth %>%
  select(AID, H4OD1Y, IYEAR4, BIO_SEX4, H4ED2, H4EC18, H4TO35, H4TO36, 
         H4GH8, H4DA23, H4TO5, H4BMICLS) %>%
  mutate(Age=IYEAR4 - H4OD1Y, 
         Obese = case_when(H4BMICLS<=3 ~ 0,
                          H4BMICLS>=4 ~ 1),
         Edu_cat = case_when(H4ED2 <=3 ~ 1,
                             H4ED2 <=5 & H4ED2 >3 ~ 2,
                             H4ED2 <=7 & H4ED2 >5 ~ 3,
                             H4ED2 <=13 & H4ED2 >7 ~ 4),
         Smoking = case_when(H4TO5==0 ~ 0,
                             H4TO5>=1 ~ 1),
         Drinking = case_when(BIO_SEX4==1 & H4TO35<=4 & H4TO35>0 & H4TO36<=4 ~ 0,
                              BIO_SEX4==1 & H4TO35>=5 & H4TO36<=2 ~ 0,
                              BIO_SEX4==1 & H4TO35<5 & H4TO36>=5 ~ 1,
                              BIO_SEX4==1 & H4TO35>=5 & H4TO36>4 ~ 1 ,
                              BIO_SEX4==1 & H4TO35>=5 & H4TO36==3 & H4TO36==4 ~ 1,
                              BIO_SEX4==2 & H4TO35<=4 & H4TO35>0 & H4TO36<=3 ~ 0,
                              BIO_SEX4==2 & H4TO35>=5 & H4TO36==1 ~ 0,
                              BIO_SEX4==2 & H4TO35<5 & H4TO36>=4 ~ 1,
                              BIO_SEX4==2 & H4TO35>=5 & H4TO36>=4 ~ 1,
                              BIO_SEX4==2 & H4TO35>=5 & H4TO36==2 & H4TO36==3 ~ 1),
         Edu_cat=factor(Edu_cat, 1:4, c("High School", "Vocation", "College", "Higher Education")),
         Smoking=factor(Smoking, 0:1, c("No", "Yes")),
         Drinking=factor(Drinking, 0:1 , c("Moderate", "Excessive")),
         Obese=factor(Obese, 0:1, c("No", "Yes")), 
         Sex=factor(BIO_SEX4, 1:2, c("Male","Female")),
         Program=factor(H4EC18, 0:1, c("No", "Yes")), 
         Fastfood=H4GH8,
         Playing_cgames=H4DA23)
  
Final_Project <- na.omit(Final_Project) #getting complete observations for all
#variables in the model

#double-checking the recoding for Obese variable

table(as_label(Final_Project$H4BMICLS))
table(as_label(Final_Project$Obese))

#double-checking the recoding for Edu_cat variable
table(as_label(Final_Project$H4ED2))
table(as_label(Final_Project$Edu_cat))

#double-checking the recoding for Drinking variable

Male <- filter(Final_Project, BIO_SEX4==1) #filter the data by male
table(Male$H4TO35, Male$H4TO36) #get frequency table for male only
Female <- filter(Final_Project, BIO_SEX4==2) #filter the data by female
table(Female$H4TO35,Female$H4TO36) #get frequency table for female only
table(as_label(Male$Drinking)) #compare results of the new variable with the old variables
table(as_label(Female$Drinking))

#double-checking the recoding for Smoking variable
table(as_label(Final_Project$H4TO5))
table(as_label(Final_Project$Smoking))

md.pattern(Final_Project)
```


##Descriptive statistics

###Drinking
```{r}
CrossTable(Final_Project$Drinking, Final_Project$Obese,
           format = c("SPSS"))
```


###Sex
```{r}
CrossTable(Final_Project$Sex, Final_Project$Obese, format = c("SPSS"))
```
###Age
```{r}
aggregate(Final_Project$Age ~ Final_Project$Obese, FUN=mean)
aggregate(Final_Project$Age ~ Final_Project$Obese, FUN=sd)
```
###Welfare recipients
```{r}
CrossTable(Final_Project$Program, Final_Project$Obese, format = c("SPSS"))
```
###Education Level
```{r}
CrossTable(Final_Project$Edu_cat, Final_Project$Obese, format = c("SPSS"))
```
###Smoking
```{r}
CrossTable(Final_Project$Smoking, Final_Project$Obese, format = c("SPSS"))
```

###Fast food
```{r}
aggregate(Final_Project$Fastfood ~ Final_Project$Obese, FUN=mean)
aggregate(Final_Project$Fastfood ~ Final_Project$Obese, FUN=sd)

```

###Playing computer
```{r}
aggregate(Final_Project$Playing_cgames ~ Final_Project$Obese, FUN=mean)
aggregate(Final_Project$Playing_cgames ~ Final_Project$Obese, FUN=sd)
```

##Checking linearity assumption using boxTidwell for continuous variables
```{r}
#linearity
Age_times_age <- Final_Project$Age * log(Final_Project$Age)#create term to test linearity

boxTidwell1 <- glm(Obese ~ Age + Age_times_age, data=Final_Project, family="binomial") #Box Tidwell technique, test the assumption of linearity

summary(boxTidwell1)

Fastfood_times_fastfood <- Final_Project$Fastfood * log(Final_Project$Fastfood) 

boxTidwell2 <- glm(Obese ~ Fastfood + Fastfood_times_fastfood, data=Final_Project, family="binomial")

summary(boxTidwell2)

Playing_times_playing <- Final_Project$Playing_cgames * log(Final_Project$Playing_cgames)

boxTidwell3 <- glm(Obese ~ Playing_cgames + Playing_times_playing, data=Final_Project, family = "binomial")

summary(boxTidwell3)

```
###Since the results from the box tidwell were not significant, the assumptions 
###for linearity were met for the continuous variables. 



##Running binary logistic regression with just drinking
```{r}
#Run a binary logistic regression 
model_bi <- glm(Obese ~ Drinking, data=Final_Project,family=binomial())
summary(model_bi)

#Get the odds ratio and 95%CIs
exp(cbind("Odds ratio" = coef(model_bi), confint(model_bi, level = 0.95)))

```

##Running the logistic model with covariates
```{r}
model1_bi <- glm(Obese ~ Drinking + Age + Sex + Edu_cat + Program + Fastfood + 
                   Playing_cgames + Smoking, data=Final_Project,family=binomial())

summary(model1_bi)
exp(cbind("Odds ratio" = coef(model1_bi), confint(model1_bi, level = 0.95)))
```

##Comparing the full and the model with just drinking variable
```{r}
logLik(model1_bi)
logLik(model_bi) #lower absolute value for the log likelihood is the better model.

lrtest(model_bi, model1_bi)
```

##Sensitivy and Specificity
```{r}
#check percent correctly predicted (example of how to do this) for bmi continuous
xt <- addmargins(table(round(predict(model1_bi, type="response")), model1_bi$model$Obese))
  xt #Note the Gold standard (reporting by participant) is the column variable and the model prediction is the row variable. 

#Sensitivity
  39/1183
  
#Specificity
  2071/2101

#Total Predicted Correctly
  (2071+39)/3284
  
```

##Looking at influence observations
```{r}
#Cook's D plot
plot(model1_bi, which=4, id.n=5, col="red", cex.id=0.60) 

#identify observations with a Cook's D greater than 0.0015 
y<-as.data.frame(cooks.distance(model1_bi))
colnames(y)[1]<-"CD"
y$obs_no<-rownames(y)
z<-y[which(y$CD>0.0015),]
z$obs_no
```


##Looking at multicollinearity
```{r}
#Variance Inflation Factors
vif(model1_bi)
```
###VIFs look good for all the variables

##Excluding influential observations
```{r}
#dropping obs with CD>0.0015
model1_bi_excluded <- update(model1_bi,subset=c(-230, -1043, -1173, -1355, -3227))
#compare coefficients between models with and without influential observations, #caveat model number of observations is not the same
compareCoefs(model1_bi, model1_bi_excluded) 
```
##The SEs with and without the influential variables were similar to each other. 

